---
title: "Plots for intro lecture to Global Change Biology"
output: html_document
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
toc_depth: 3
number_sections: true
---

# Box models

```{r}
library(ggplot2)
library(tidyr)
library(gganimate)
library(gifski)
library(purrr)
library(stringr)
```

This produces plots and animations for the intro to biogeochemical cycles as part of the lecture *Global Change Biology* (ETH Zürich).

## Function definition

Illustrations are made for the dynamics of a 1-box model with first-order decay. Such a model is defined as a function below.

```{r}
## //////////////////////////////////////////////////
## 1-BOX MODEL
## --------------------------------------------------
onebox <- function( c_influx, tau, cpool0 ) {
  ## ------------------------------------------------
  ## c_influx:  flux into pool (GtC a-1)
  ## cpool0:    initial pool size (GtC)
  ## tau:       turnover (residence) time (a)
  ## ------------------------------------------------
  tauisvec <- ifelse(length(tau)==length(c_influx), TRUE, FALSE)

  ## determine integration length (number of time steps) from length of 'c_influx'
  len <- length(c_influx)

  ## initialise output variable (time series of pool size)
  out_cpool <- rep( NA, len )

  ## copy initial pool size to first element in output time series
  cpool <- cpool0

  ## integrate over each time step (this is an implementation of the differential equation)
  for (yr in seq(len) ) {

    ## copy current pool size to output time series
    out_cpool[yr] <- cpool

    ## update pool size with input and decay
    cpool <- cpool + c_influx[yr] - 1/ifelse(tauisvec, tau[yr], tau) * cpool

  }

  ## function return value is a vector containing the output time series
  return( out_cpool )

}
```

## Pulse-decay

```{r}
nt       <- 400
cpool0    <- 3000
c_influx  <- rep( 0, nt )
c_influx[50] <- 100
tau       <- 50

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = 50, cpool0 = 0))

gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  facet_wrap( ~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")) ) +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/gganim_1box_pulse.gif", loop = TRUE))
```

## Spin up with two turnover rates

```{r}
c_influx  <- rep( 60, nt )

df <- tibble(year = seq(nt), c_influx_tau50 = c_influx, c_influx_tau60 = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx_tau50, tau = 50, cpool0 = 0),
                c_pool_tau60 = onebox(c_influx = c_influx_tau60, tau = 60, cpool0 = 0))


gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  mutate(tau = purrr::map_chr(var, ~ifelse(str_detect(., pattern = "_tau50$"), "50", "60"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau50"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau60"))) %>%
  ggplot(aes(x = year, y = c_, color = tau)) +
  geom_line(size = 1) +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/gganim_1box_spinup.gif", loop = TRUE))
```

## Equilibrium?

```{r}
c_influx  <- rep( 60, nt ) + rnorm(n = nt, mean = 0, sd = 5)

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx, tau = 50, cpool0 = 0))

gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/gganim_1box_steadystate.gif", loop = TRUE))
```

## Step change in input flux

```{r}
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx_tau50 = c_influx, c_influx_tau60 = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx_tau50, tau = 50, cpool0 = 3000),
                c_pool_tau60 = onebox(c_influx = c_influx_tau60, tau = 60, cpool0 = 3000*60/50))

gganim <- df %>%
  mutate(c_pool_tau50 = c_pool_tau50 - c_pool_tau50[1], c_pool_tau60 = c_pool_tau60 - c_pool_tau60[1]) %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  mutate(tau = purrr::map_chr(var, ~ifelse(str_detect(., pattern = "_tau50$"), "50", "60"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau50"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau60"))) %>%
  ggplot(aes(x = year, y = c_, color = tau)) +
  geom_line(size = 1) +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "Increase in C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/gganim_1box_step.gif", loop = TRUE))
```

## Step change in turnover time

```{r}
c_influx  <- rep( 60, nt )
vec_tau <-  c(rep( 50, 100 ), rep( 60, (nt-100) ))

df <- tibble(year = seq(nt),
             c_influx = c_influx,
             tau = vec_tau) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = vec_tau, cpool0 = 3000)) %>%
  dplyr::mutate(c_uptake = c_pool - lag(c_pool, 1))

gganim <- df %>%
  pivot_longer(c(tau, c_pool, c_uptake), names_to = "var", values_to = "c_") %>%
  mutate(var = factor(var, levels = c("tau", "c_pool", "c_uptake"))) %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  facet_wrap(~var, nrow = 3, labeller = labeller(var = c("tau" = "tau (yr)", "c_pool" = "C pool (Gt C)", "c_uptake" = "C uptake (Gt C/yr)")), scales = "free") +
  labs(y = "", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/gganim_1box_tauincrease.gif", loop = TRUE))
```

# Maps

Some maps are created in 

## Photosynthesis

```{r}
## load data created in eval_pmodel/eval_pmodel_global.Rmd
load("~/eval_pmodel/data/nc_gpp_mean.Rdata")

plot_map3(
  nc_gpp_mean, 
  colorscale = viridis, 
  nbin = 15, 
  # plot_title = "GPP: Mean of other models", 
  plot_subtitle = expression( paste("g C m"^-2, " yr"^-1 ) ), 
  breaks = c(seq(0, 3000, 200), Inf)
  )
ggsave("~/teaching/gcb/photosynthesis.pdf", width = 8, height = 4)
```

## Land use

```{r}
nc_lu <- read_nc_onefile("~/data/landuse_data/hyde32_gcp2017/landuse_hyde32_gcp2017_halfdeg.cdf") %>%    
  slice_nc(75) %>% 
  extract_nc(var = "crop")

nc_lu$vars$crop <- ifelse(nc_lu$vars$crop > 1.0, 1.0, nc_lu$vars$crop)
nc_lu$vars$crop <- ifelse(nc_lu$vars$crop < 0.0, 0.0, nc_lu$vars$crop)

plot_map3(
  nc_lu, 
  colorscale = c("wheat3","green4","yellow","orange","red3"), 
  nbin = 15, 
  plot_title = "Cropland",
  plot_subtitle = "HYDE 3.2, area fraction", 
  breaks = seq(0, 1, 0.1)
  # legend_title = "(unitless)"
  )
ggsave("~/teaching/gcb/cropland.pdf", width = 8, height = 4)

```

## P/PET map

```{r}
require(ncdf4)
library(ggplot2)
library(rbeni)

ncfiln <- "~/data/greve/ep_over_p_cru_ncep.nc"
if (!file.exists(ncfiln)) {
  epop <- array( 1, dim=c(720,360) )
} else {
  nc <- nc_open( ncfiln )
  epop <- ncvar_get( nc, varid="EP_OVER_P_CRU_NCEP" )
}

gg <- plot_map3(
  1/epop,
  breaks = c(0,0.2,0.3,0.4,0.5,0.7,1,1.3,1.6,2,2.5,3,Inf),
  colorscale = c( "royalblue3", "wheat", "tomato" ) %>% rev(),
  combine = FALSE,
  legend_title = "P/PET"
)

cowplot::plot_grid(gg$ggmap, gg$gglegend, ncol = 2, rel_widths = c(1, 0.2))

ggsave("~/teaching/gcb/map_p_pet.pdf", width = 8, height = 4)
```

## fAPAR map

```{r}
nc_fapar <- read_nc_onefile("~/data/modis_lai_fpar_zmaw/MODIS-C006_MOD15A2__LAI_FPAR__LPDAAC__GLOBAL_0.5degree__UHAM-ICDC__2000_2018__MON__fv0.02_MEAN.nc") %>% 
  extract_nc("fpar")

gg <- plot_map3(
  nc_fapar,
  colorscale = c( "wheat", "darkgreen" ), #
  combine = TRUE
)
gg
ggsave("~/teaching/gcb/map_fapar.pdf", width = 8, height = 4)
```

## Photosynthesis FvCB model

```{r}
library(rpmodel)
library(dplyr)
# modified seq() function to get a logarithmically spaced sequence
lseq <- function(from=1, to=100000, length.out=6) {
  exp(seq(log(from), log(to), length.out = length.out))
}

## Set parameters
beta <- 146          # unit cost ratio a/b
c_cost <- 0.41
gamma <- 0.105       # unit cost ratio c/b
kphio <- 0.05        # quantum yield efficiency
c_molmass <- 12.0107 # molar mass, g / mol

## Define environmental conditions
tc <- 20             # temperature, deg C
ppfd <- 100/(24*60*60)*1e6         # mol/m2/d
vpd  <- 1000         # Pa
co2  <- 400          # ppm
elv  <- 0            # m.a.s.l.
fapar <- 1           # fraction
patm <- 101325       # Pa

out_pmodel <- rpmodel::rpmodel(
  tc             = tc,
  vpd            = vpd,
  co2            = co2,
  elv            = elv,
  kphio          = kphio,
  beta           = beta,
  fapar          = fapar,
  ppfd           = ppfd,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE
  )

calc_aj <- function(ci, gammastar, kphio, ppfd){
  kphio * ppfd * (ci - gammastar)/(ci + 2 * gammastar)
}
calc_ac <- function(ci, gammastar, kmm, vcmax){
  vcmax * (ci - gammastar)/(ci + kmm)
}

co2_to_ca <- function( co2, patm ){
  ( 1.0e-6 ) * co2 * patm         # Pa, atms. CO2
}

calc_a_gs <- function(ci, gs, ca){
  gs * (ca - ci)
}

df_ci <- tibble(
  ci = seq(0, 1000, length.out = 100)) %>% 
  rowwise() %>% 
  mutate(ci_pa = co2_to_ca(ci, patm = patm)) %>% 
  mutate(a_j = calc_aj(ci_pa, out_pmodel$gammastar, kphio = kphio, ppfd = ppfd)) %>% 
  mutate(a_c = calc_ac(ci_pa, out_pmodel$gammastar, out_pmodel$kmm, vcmax = out_pmodel$vcmax)) %>% 
  mutate(a_act = min(a_j, a_c)) %>% 
  mutate(a_gs = calc_a_gs(ci_pa, gs = out_pmodel$gs, ca = out_pmodel$ca))
  
df_ci %>% 
  pivot_longer(cols = c(a_j, a_c, a_gs), names_to = "Rate", values_to = "a_") %>% 
  ggplot(aes(x = ci)) +
  geom_line(aes(y = a_, color = Rate)) +
  geom_line(aes(y = a_act)) +
  # geom_line(aes(y = a_gs), linetype = "dashed") +
  xlim(50, 1000) + ylim(0, 80) +
  labs(x = expression(paste("c"[i], " (ppm)")), y = expression(paste("A (", mu, "mol m" ^{-2}," s" ^{-1}, ")"))) +
  scale_color_discrete(
    name = "Rate", 
    breaks = c("a_gs", "a_j", "a_c"),
    labels = c(expression(paste("A"[gs])), expression(paste("A"[J])), expression(paste("A"[C])))
    )

ggsave("~/teaching/gcb/assimilation.pdf", width = 4, height = 2)
```

