---
title: "Plots for intro lecture to Global Change Biology"
output: html_document
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
toc_depth: 3
number_sections: true
---

# Box models

```{r}
library(ggplot2)
library(tidyr)
library(gganimate)
library(gifski)
library(purrr)
library(stringr)
```

This produces plots and animations for the intro to biogeochemical cycles as part of the lecture *Global Change Biology* (ETH Zürich).

## Function definition

Illustrations are made for the dynamics of a 1-box model with first-order decay. Such a model is defined as a function below.

### 1-box
```{r}
## //////////////////////////////////////////////////
## 1-BOX MODEL
## --------------------------------------------------
onebox <- function( c_influx, tau, cpool0, return_cpool = TRUE) {
  ## ------------------------------------------------
  ## c_influx:  flux into pool (GtC a-1)
  ## cpool0:    initial pool size (GtC)
  ## tau:       turnover (residence) time (a)
  ## ------------------------------------------------
  tauisvec <- ifelse(length(tau)==length(c_influx), TRUE, FALSE)

  ## determine integration length (number of time steps) from length of 'c_influx'
  len <- length(c_influx)

  ## initialise output variable (time series of pool size)
  out_cpool <- rep( NA, len )
  out_outflux <- rep(NA, len)
  
  ## copy initial pool size to first element in output time series
  cpool <- cpool0

  ## integrate over each time step (this is an implementation of the differential equation)
  for (yr in seq(len) ) {

    ## copy current pool size to output time series
    out_cpool[yr] <- cpool
    
    ## outflux
    outflux <- 1/ifelse(tauisvec, tau[yr], tau) * cpool

    ## update pool size with input and decay
    cpool <- cpool + c_influx[yr] - outflux
    out_outflux[yr] <- outflux

  }

  ## function return value is a vector containing the output time series
  if (return_cpool){
    return( out_cpool )
  } else {
    return( out_outflux )
  }

}
```

### 1-box non-linear in tau
```{r}
## //////////////////////////////////////////////////
## 1-BOX MODEL
## --------------------------------------------------
onebox_nonlintau <- function(df, cpool0) {
  ## ------------------------------------------------
  ## c_influx:  flux into pool (GtC a-1)
  ## cpool0:    initial pool size (GtC)
  ## tau:       turnover (residence) time (a)
  ## ------------------------------------------------
  ## determine integration length (number of time steps) from length of 'c_influx'
  len <- length(df$c_influx)

  ## initialise output variable (time series of pool size)
  df$cpool <- rep( NA, len )

  ## copy initial pool size to first element in output time series
  df$cpool[1] <- cpool0
  df <- df %>% mutate(tau = NA)
  df$tau[1] <- 50

  ## integrate over each time step (this is an implementation of the differential equation)
  for (yr in 2:len ) {
    
    df$tau[yr] <- (50 - (df$cpool[yr-1] - 3000)/100)
    
    ## outflux
    outflux <- 1/df$tau[yr] * df$cpool[yr-1]

    ## update pool size with input and decay
    df$cpool[yr] <- df$cpool[yr-1] + c_influx[yr] - outflux

  }

  ## function return value is a vector containing the output time series
  return( df )

}
```

### 2-box model
```{r}
## //////////////////////////////////////////////////
## 1-BOX MODEL
## --------------------------------------------------
twobox <- function( df, a1, tau1, tau2, cpool0_1, cpool0_2, return_cpool = TRUE) {
  ## ------------------------------------------------
  ## c_influx:  flux into pool (GtC a-1)
  ## cpool0:    initial pool size (GtC)
  ## tau:       turnover (residence) time (a)
  ## ------------------------------------------------
  tauisvec <- ifelse(length(tau1)==length(df$c_influx), TRUE, FALSE)

  ## determine integration length (number of time steps) from length of 'c_influx'
  len <- nrow(df)

  ## initialise output variable (time series of pool size)
  df$cpool1 <- rep( NA, len )
  df$cpool2 <- rep( NA, len )

  ## copy initial pool size to first element in output time series
  df$cpool1[1] <- cpool0_1
  df$cpool2[1] <- cpool0_2

  ## integrate over each time step (this is an implementation of the differential equation)
  for (yr in 2:len ) {
    
    ## outflux
    outflux1 <- 1/ifelse(tauisvec, tau1[yr], tau1) * df$cpool1[yr-1]
    outflux2 <- 1/ifelse(tauisvec, tau2[yr], tau2) * df$cpool2[yr-1]

    ## update pool size with input and decay
    df$cpool1[yr] <- df$cpool1[yr-1] + a1 * df$c_influx[yr] - outflux1
    df$cpool2[yr] <- df$cpool2[yr-1] + (1 - a1) * df$c_influx[yr] - outflux2

  }

  ## function return value is a vector containing the output time series
  return( df )

}
```

### 2-box non-linear in allocation
```{r}
## //////////////////////////////////////////////////
## 1-BOX MODEL
## --------------------------------------------------
twobox_nonlin <- function( df, a1, tau1, tau2, cpool0_1, cpool0_2, return_cpool = TRUE) {
  ## ------------------------------------------------
  ## c_influx:  flux into pool (GtC a-1)
  ## cpool0:    initial pool size (GtC)
  ## tau:       turnover (residence) time (a)
  ## ------------------------------------------------
  ## determine integration length (number of time steps) from length of 'c_influx'
  len <- nrow(df)

  ## initialise output variable (time series of pool size)
  df$cpool1 <- rep( NA, len )
  df$cpool2 <- rep( NA, len )
  df$a1 <- rep( NA, len )

  ## copy initial pool size to first element in output time series
  df$cpool1[1] <- cpool0_1
  df$cpool2[1] <- cpool0_2
  df$a1[1] <- 0.75

  ## integrate over each time step (this is an implementation of the differential equation)
  for (yr in 2:len ) {
    
    df$a1[yr] <- 0.25 + (df$c_influx[yr] - 60)/50
    
    ## outflux
    outflux1 <- 1/tau1 * df$cpool1[yr-1]
    outflux2 <- 1/tau2 * df$cpool2[yr-1]

    ## update pool size with input and decay
    df$cpool1[yr] <- df$cpool1[yr-1] + df$a1[yr] * df$c_influx[yr] - outflux1
    df$cpool2[yr] <- df$cpool2[yr-1] + (1 - df$a1[yr]) * df$c_influx[yr] - outflux2

  }

  ## function return value is a vector containing the output time series
  return( df )

}
```

## Pulse-decay

```{r}
nt       <- 400
cpool0    <- 3000
c_influx  <- rep( 0, nt )
c_influx[50] <- 100
tau       <- 50

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = 50, cpool0 = 0))

gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  facet_wrap( ~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")) ) +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_pulse.gif", loop = TRUE))
```

## Spin up with one turnover rate

```{r}
c_influx  <- rep( 60, nt )

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = 50, cpool0 = 0),
                c_outflux = onebox(c_influx = c_influx, tau = 50, cpool0 = 0, return_cpool = FALSE))

gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "red") +
  facet_wrap(~var, nrow = 3, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool", "c_outflux" = "C outflux")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_spinup.gif", loop = TRUE))
```

## Spin up with two turnover rates

```{r}
c_influx  <- rep( 60, nt )

df <- tibble(year = seq(nt), c_influx_tau50 = c_influx, c_influx_tau60 = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx_tau50, tau = 50, cpool0 = 0),
                c_pool_tau60 = onebox(c_influx = c_influx_tau60, tau = 60, cpool0 = 0))


gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  mutate(tau = purrr::map_chr(var, ~ifelse(str_detect(., pattern = "_tau50$"), "50", "60"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau50"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau60"))) %>%
  ggplot(aes(x = year, y = c_, color = tau)) +
  geom_line(size = 1) +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_spinup.gif", loop = TRUE))
```

## Equilibrium?

```{r}
c_influx  <- rep( 60, nt ) + rnorm(n = nt, mean = 0, sd = 5)

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx, tau = 50, cpool0 = 0))

gganim <- df %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_steadystate.gif", loop = TRUE))
```

## Step change in input flux, one turnovover time

```{r}
nt <- 400
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = 50, cpool0 = 3000))

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$c_pool[2] - df2$c_pool[1])/df2$c_pool[1])

df4 <- tibble(x = (log(df2$c_influx[2]) - log(df2$c_influx[1])),
              y =  (log(df2$c_pool[2]) - log(df2$c_pool[1])))

df3 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln.pdf", width = 4, height = 3.5)

gganim <- df %>%
  mutate(c_pool = c_pool) %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "red") +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_step.gif", loop = TRUE))
```

## Step change in input flux, two turnovover times

```{r}
nt <- 1000
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  dplyr::mutate(c_pool_tau50 = onebox(c_influx = c_influx, tau = 50, cpool0 = 3000),
                c_pool_tau60 = onebox(c_influx = c_influx, tau = 60, cpool0 = 3000*60/50))

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx_tau50[2] - df2$c_influx_tau50[1])/df2$c_influx_tau50[1],
              y_tau50 =  (df2$c_pool_tau50[2] - df2$c_pool_tau50[1])/df2$c_pool_tau50[1],
              y_tau60 =  (df2$c_pool_tau60[2] - df2$c_pool_tau60[1])/df2$c_pool_tau60[1]) %>% 
  pivot_longer(cols = starts_with("y_tau"), names_to = "turnover", values_to = "y", names_prefix = "y_tau")

df3 %>% 
  ggplot(aes(x, y, color = as.factor(turnover), size = 100 - turnover/10)) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_2turnover.pdf", width = 5, height = 3.5)

gganim <- df %>%
  mutate(c_pool_tau50 = c_pool_tau50 - c_pool_tau50[1], c_pool_tau60 = c_pool_tau60 - c_pool_tau60[1]) %>%
  pivot_longer(starts_with("c_"), names_to = "var", values_to = "c_") %>%
  mutate(tau = purrr::map_chr(var, ~ifelse(str_detect(., pattern = "_tau50$"), "50", "60"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau50"))) %>%
  mutate(var = purrr::map_chr(var, ~str_remove(., "_tau60"))) %>%
  ggplot(aes(x = year, y = c_, color = tau)) +
  geom_line(size = 1) +
  facet_wrap(~var, nrow = 2, labeller = labeller(var = c("c_influx" = "C influx", "c_pool" = "Increase in C pool")), scales = "free") +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_step_2turnover.gif", loop = TRUE))
```

## Step change in turnover time

```{r}
c_influx  <- rep( 60, nt )
vec_tau <-  c(rep( 50, 100 ), rep( 60, (nt-100) ))

df <- tibble(year = seq(nt),
             c_influx = c_influx,
             tau = vec_tau) %>%
  dplyr::mutate(c_pool = onebox(c_influx = c_influx, tau = vec_tau, cpool0 = 3000)) %>%
  dplyr::mutate(c_uptake = c_pool - lag(c_pool, 1))

gganim <- df %>%
  pivot_longer(c(tau, c_pool, c_uptake), names_to = "var", values_to = "c_") %>%
  mutate(var = factor(var, levels = c("tau", "c_pool", "c_uptake"))) %>%
  ggplot(aes(x = year, y = c_)) +
  geom_line(size = 1, color = "tomato") +
  facet_wrap(~var, nrow = 3, labeller = labeller(var = c("tau" = "tau (yr)", "c_pool" = "C pool (Gt C)", "c_uptake" = "C uptake (Gt C/yr)")), scales = "free") +
  labs(y = "", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_tauincrease.gif", loop = TRUE))
```

## Non-linearity in turnover time

```{r}
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt),
             c_influx = c_influx) %>%
  onebox_nonlintau(cpool0 = 3000)

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpool[2] - df2$cpool[1])/df2$cpool[1])

df3 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_nonlin_turnover.pdf", width = 5, height = 3.5)
```

## 2-box model, step change

```{r}
nt <- 400
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  twobox(tau1 = 30, tau2 = 60, cpool0_1 = 0.5*3000, cpool0_2 = 0.5*3000*30/50, a1 = 0.5) %>% 
  mutate(cpooltot = cpool1 + cpool2)

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpool1[2] - df2$cpool1[1])/df2$cpool1[1])

df4 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpool2[2] - df2$cpool2[1])/df2$cpool2[1])

df5 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpooltot[2] - df2$cpooltot[1])/df2$cpooltot[1])

df3 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_cpool1.pdf", width = 3, height = 2)

df4 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_cpool2.pdf", width = 3, height = 2)

## individual pools
gganim <- df %>%
  pivot_longer(cols = c(cpool1, cpool2), names_to = "var", values_to = "cpool") %>%
  ggplot(aes(x = year, y = cpool, color = var)) +
  geom_line(size = 1) +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_step_2pools.gif", loop = TRUE))

## total
gganim <- df %>%
  select(year, cpooltot) %>%
  ggplot(aes(x = year, y = cpooltot)) +
  geom_line(size = 1) +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_step_tot.gif", loop = TRUE))
```

## 2-box model, step change, allocation

```{r}
nt <- 400
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  twobox(tau1 = 30, tau2 = 60, cpool0_1 = 0.5*60*30, cpool0_2 = 0.5*60*60, a1 = 0.5) %>% 
  mutate(cpooltot1 = cpool1 + cpool2) %>%
  twobox(tau1 = 30, tau2 = 60, cpool0_1 = 0.75*60*30, cpool0_2 = 0.25*60*60, a1 = 0.75) %>% 
  mutate(cpooltot2 = cpool1 + cpool2)

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpooltot1[2] - df2$cpooltot1[1])/df2$cpooltot1[1])

df4 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpooltot2[2] - df2$cpooltot2[1])/df2$cpooltot2[1])

df3 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_cpool1.pdf", width = 3, height = 2)

df4 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_cpool2.pdf", width = 3, height = 2)

## total pools
gganim <- df %>%
  pivot_longer(cols = c(cpooltot1, cpooltot2), names_to = "var", values_to = "cpooltot") %>%
  ggplot(aes(x = year, y = cpooltot, color = var)) +
  geom_line(size = 1) +
  labs(y = "Carbon (Gt C)", x = "Year") +
  theme(strip.text = element_text(face = "bold")) +
  theme_grey(base_size = 20) +
  transition_reveal(year)

animate(gganim, nframes = 100, width = 700, height = 500, fps = 10, renderer = gifski_renderer("~/teaching/gcb/fig/gganim_1box_step_2pools_allocation.gif", loop = TRUE))
```

## Non-linearity in allocation

```{r}
nt <- 400
c_influx  <- c(rep( 60, 100 ), rep( 80, (nt-100) ))

df <- tibble(year = seq(nt), c_influx = c_influx) %>%
  twobox_nonlin(tau1 = 30, tau2 = 60, cpool0_1 = 0.5*60*30, cpool0_2 = 0.5*60*60, a1 = 0.5) %>% 
  mutate(cpooltot1 = cpool1 + cpool2)

df2 <- slice(df, 1) %>% 
  bind_rows(slice(df, nrow(df)))

df3 <- tibble(x = (df2$c_influx[2] - df2$c_influx[1])/df2$c_influx[1],
              y =  (df2$cpooltot1[2] - df2$cpooltot1[1])/df2$cpooltot1[1])

df3 %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  theme_classic() +
  xlim(0,1) + ylim(0,1) +
  labs(x = "dI/I", y = "dC*/C*")
ggsave("~/teaching/gcb/fig/dln_cpool_nonlinalloc.pdf", width = 3, height = 2)
```

# Maps

Some maps are created in 

## Photosynthesis

```{r}
## load data created in eval_pmodel/eval_pmodel_global.Rmd
load("~/eval_pmodel/data/nc_gpp_mean.Rdata")

plot_map3(
  nc_gpp_mean, 
  colorscale = viridis, 
  nbin = 15, 
  # plot_title = "GPP: Mean of other models", 
  plot_subtitle = expression( paste("g C m"^-2, " yr"^-1 ) ), 
  breaks = c(seq(0, 3000, 200), Inf)
  )
ggsave("~/teaching/gcb/fig/photosynthesis.pdf", width = 8, height = 4)
```

## Land use

```{r}
nc_lu <- read_nc_onefile("~/data/landuse_data/hyde32_gcp2017/landuse_hyde32_gcp2017_halfdeg.cdf") %>%    
  slice_nc(75) %>% 
  extract_nc(var = "crop")

nc_lu$vars$crop <- ifelse(nc_lu$vars$crop > 1.0, 1.0, nc_lu$vars$crop)
nc_lu$vars$crop <- ifelse(nc_lu$vars$crop < 0.0, 0.0, nc_lu$vars$crop)

plot_map3(
  nc_lu, 
  colorscale = c("wheat3","green4","yellow","orange","red3"), 
  nbin = 15, 
  plot_title = "Cropland",
  plot_subtitle = "HYDE 3.2, area fraction", 
  breaks = seq(0, 1, 0.1)
  # legend_title = "(unitless)"
  )
ggsave("~/teaching/gcb/fig/cropland.pdf", width = 8, height = 4)

```

## Land use trend

```{r}
nc_lu <- nc_to_df("~/data/landuse_data/hyde32_gcp2017/landuse_hyde32_gcp2017_halfdeg_1991_2017_diff.nc", varnam = "crop")

plot_map3(
  nc_lu, 
  varnam = "crop",
  colorscale = rev(c( "royalblue4", "royalblue2", "wheat", "tomato2", "tomato4" )), 
  nbin = 15, 
  plot_title = "Cropland change",
  plot_subtitle = "HYDE 3.2, change in area fraction, 1970-2017", 
  breaks = seq(-0.5,  0.5, 0.1),
  latmin = -60, latmax = 80
  # legend_title = "(unitless)"
  )
ggsave("~/teaching/gcb/fig/cropland_diff_diff_1970_2017.pdf", width = 8, height = 4)

```


## P/PET map

```{r}
require(ncdf4)
library(ggplot2)
library(rbeni)
library(tidyverse)

ncfiln <- "~/data/greve/ep_over_p_cru_ncep.nc"
if (!file.exists(ncfiln)) {
  epop <- array( 1, dim=c(720,360) )
} else {
  nc <- nc_open( ncfiln )
  epop <- ncvar_get( nc, varid="EP_OVER_P_CRU_NCEP" )
}

gg <- plot_map3(
  1/epop,
  breaks = c(0,0.2,0.3,0.4,0.5,0.7,1,1.3,1.6,2,2.5,3,Inf),
  colorscale = c( "royalblue3", "wheat", "tomato" ) %>% rev(),
  combine = FALSE,
  legend_title = "P/PET"
  )

gg$ggmap <- gg$ggmap +
  ylim(-60, 85)

cowplot::plot_grid(gg$ggmap, gg$gglegend, ncol = 2, rel_widths = c(1, 0.2))

ggsave("~/teaching/gcb/fig/map_p_pet.pdf", width = 8, height = 4)
```

## fAPAR map

```{r}
nc_fapar <- read_nc_onefile("~/data/modis_lai_fpar_zmaw/MODIS-C006_MOD15A2__LAI_FPAR__LPDAAC__GLOBAL_0.5degree__UHAM-ICDC__2000_2018__MON__fv0.02_MEAN.nc") %>% 
  extract_nc("fpar")

gg <- plot_map3(
  nc_fapar,
  colorscale = c( "wheat", "darkgreen" ), #
  combine = TRUE
)
gg
ggsave("~/teaching/gcb/fig/map_fapar.pdf", width = 8, height = 4)
```

## LAI map

```{r}
library(rbeni)
library(tidyverse)

# df_lai <- nc_to_df("~/data/modis_lai_fpar_zmaw/MODIS-C006_MOD15A2__LAI_FPAR__LPDAAC__GLOBAL_0.5degree__UHAM-ICDC__2000_2018__MON__fv0.02_MEAN.nc", varnam = "lai")

gg <- plot_map3(
  df_lai,
  varnam = "lai",
  colorscale = c( "wheat", "darkgreen" ), #
  combine = TRUE,
  breaks = c(0, 0.1, 0.2, 0.5, 1, 1.5, 2, 3, 4, 5, 6), 
  latmin = -55, latmax = 80
)
gg
ggsave("~/teaching/gcb/fig/map_lair.pdf", width = 8, height = 4)
```

## Photosynthesis FvCB model

```{r}
library(rpmodel)
library(dplyr)
# modified seq() function to get a logarithmically spaced sequence
lseq <- function(from=1, to=100000, length.out=6) {
  exp(seq(log(from), log(to), length.out = length.out))
}

## Set parameters
beta <- 146          # unit cost ratio a/b
c_cost <- 0.41
gamma <- 0.105       # unit cost ratio c/b
kphio <- 0.05        # quantum yield efficiency
c_molmass <- 12.0107 # molar mass, g / mol

## Define environmental conditions
tc <- 20             # temperature, deg C
ppfd <- 100/(24*60*60)*1e6         # mol/m2/d
vpd  <- 1000         # Pa
co2  <- 400          # ppm
elv  <- 0            # m.a.s.l.
fapar <- 1           # fraction
patm <- 101325       # Pa

out_pmodel <- rpmodel::rpmodel(
  tc             = tc,
  vpd            = vpd,
  co2            = co2,
  elv            = elv,
  kphio          = kphio,
  beta           = beta,
  fapar          = fapar,
  ppfd           = ppfd,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE
  )

calc_aj <- function(ci, gammastar, kphio, ppfd){
  kphio * ppfd * (ci - gammastar)/(ci + 2 * gammastar)
}
calc_ac <- function(ci, gammastar, kmm, vcmax){
  vcmax * (ci - gammastar)/(ci + kmm)
}

co2_to_ca <- function( co2, patm ){
  ( 1.0e-6 ) * co2 * patm         # Pa, atms. CO2
}

calc_a_gs <- function(ci, gs, ca){
  gs * (ca - ci)
}

df_ci <- tibble(
  ci = seq(0, 1000, length.out = 100)) %>% 
  rowwise() %>% 
  mutate(ci_pa = co2_to_ca(ci, patm = patm)) %>% 
  mutate(a_j = calc_aj(ci_pa, out_pmodel$gammastar, kphio = kphio, ppfd = ppfd)) %>% 
  mutate(a_c = calc_ac(ci_pa, out_pmodel$gammastar, out_pmodel$kmm, vcmax = out_pmodel$vcmax)) %>% 
  mutate(a_act = min(a_j, a_c)) %>% 
  mutate(a_gs = calc_a_gs(ci_pa, gs = out_pmodel$gs, ca = out_pmodel$ca))
  
df_ci %>% 
  # pivot_longer(cols = c(a_j, a_c, a_gs), names_to = "Rate", values_to = "a_") %>% 
  pivot_longer(cols = c(a_j, a_c), names_to = "Rate", values_to = "a_") %>% 
  ggplot(aes(x = ci)) +
  geom_line(aes(y = a_, color = Rate)) +
  geom_line(aes(y = a_act)) +
  # geom_line(aes(y = a_gs), linetype = "dashed") +
  xlim(50, 1000) + ylim(0, 80) +
  labs(x = expression(paste("c"[i], " (ppm)")), y = expression(paste("A (", mu, "mol m" ^{-2}," s" ^{-1}, ")"))) +
  scale_color_discrete(
    name = "Rate", 
    breaks = c("a_gs", "a_j", "a_c"),
    labels = c(expression(paste("A"[gs])), expression(paste("A"[J])), expression(paste("A"[C])))
    ) +
  theme_classic()

ggsave("~/teaching/gcb/fig/assimilation.pdf", width = 5, height = 3)
```

